---
title: "1. Getting Started"
author: "Vivek H. Patil"
date: "September 17, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting and Loading Packages

## Installing Packages from CRAN

```{r eval=FALSE}
install.packages("pkgname")
install.packages("pkgname1","pkgname2") # For two or more packages 
```

## Installing a package from GitHub

```{r eval=FALSE}
if (!require("devtools")) install.packages("devtools") # Checks if the devtools package is available and if not, installs it from CRAN
devtools::install_github("githubaccountname/packagerepository")
```

## Installed packages should be loaded for them to offer their functions for use

```{r eval=FALSE}
library(pkgname)
```


# Getting data

## Load R file

```{r eval=FALSE}
load("filename.Rda")
```

## Delimited files text files
```{r eval=FALSE}
dataname=read.csv("datafilename.csv")# use ?read.csv to see options
```

## Excel Files
```{r eval=FALSE}

install.packages("readxl") # https://github.com/hadley/readxl
read_excel("excelfilename.xls") # reads xlsx extensions also
read_excel("excelfilename.xlsx", sheet = "nameofsheet")
read_excel("excelfilename.xlsx", sheet = 3)# for third sheet, as an example
read_excel("my-spreadsheet.xls", na = "NA") # if NAs are represented by non-blank cells, specify them with na=
```

## SPSS/SAS/Stata files

```{r eval=FALSE}
install.packages("haven") # https://github.com/hadley/haven
read_sas("path/to/file") # SAS file
read_por("path/to/file") # SPSS por files
read_sav("path/to/file") # SPSS sav files
read_dta("path/to/file")# Stata
```

## Many others
```{r eval=FALSE}
?read.table
?readClipboard
# Checkout the readr package from Hadley: https://github.com/hadley/readr
```

# Saving data

```{r eval=FALSE}
save(dataframename,file="filename.Rda") # To R data
write.csv(dataframename,file="myfilename.csv") # to csv...see also ?write.csv
write.table(dataframename,file="myfilename.tsv",sep="\t") # to tab-separated file...see also ?write.table
```


# Creating a dataframe

```{r}
Person=c("A","B","C","D","E")
Age=c(15,20,25,30,35)
Height=c(60,63,75,79,56)
page=data.frame(Person,Age,Height)
```


# Little knowledge of data

Number of rows and columns (in that sequence)

```{r}
dim(iris)
```

Number of rows
```{r}
nrow(iris)
```

Number of columns
```{r}
ncol(iris)
```

Names of variables
```{r}
names(iris) # colnames(iris) also gives that information
```

First 6 rows 
```{r}
head(iris)
```

First 2 rows
```{r}
head(iris,2) # alternately, can use iris[1:2,]
```

Last 6 rows
```{r}
tail(iris) # Number of rows can be controlled, see earlier example involving the head command
```

First row
```{r}
iris[1,]
```
First row, first column
```{r}
iris[1,1]
```

Name of third column
```{r}
names(iris)[3]
```

3 entries from third column
```{r}
head(iris[3],3) # alternately, we can use iris[1:3,3]
```
Structure of the dataframe
```{r}
str(iris)
```
Summary of the dataframe
```{r}
summary(iris) # for factor/categorical variables, this gives a count of all categories
```


# Make few changes to data

## Changing names of columns/headers

```{r eval=FALSE}
names(dataname)=c("newcol1","newcol2","newcol3") # for a 3 column dataset
names(dataname)[3]="latestcol3name" # changing name for only the third column
```

